Substitution rate responses to changes in effective population size.

### 2020-03-31 manuscript/manuscript.tex
Moved the notations into new-commands.tex

\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{adjustbox}
\usepackage{float}
\usepackage{caption}
\usepackage{mdframed}
\usepackage{lmodern}
\usepackage{bm,bbold}
\usepackage{xfrac, nicefrac}
\usepackage{xcolor}
\usepackage{lmodern}
\usepackage{enumitem}
\usepackage{pgfplots, pgf,tikz}
\usepackage[margin=60pt]{geometry}
\pdfinclusioncopyfonts=1

\definecolor{RED}{HTML}{EB6231}
\definecolor{YELLOW}{HTML}{E29D26}
\definecolor{BLUE}{HTML}{5D80B4}
\definecolor{LIGHTGREEN}{HTML}{6ABD9B}
\definecolor{GREEN}{HTML}{8FB03E}
\definecolor{PURPLE}{HTML}{BE1E2D}
\definecolor{BROWN}{HTML}{A97C50}
\definecolor{PINK}{HTML}{DA1C5C}
\pgfplotsset{every axis/.append style={line width=1pt}}
\pgfplotscreateplotcyclelist{colors}{LIGHTGREEN\\YELLOW\\RED\\GREEN\\BLUE\\}
\captionsetup{width=0.85\textwidth}

\newcommand{\specialcell}[2][c]{%
	\begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\DeclareMathOperator{\E}{\mathbb{E}}
\newcommand{\der}{\mathrm{d}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\indice}{l}
\newcommand{\indiceexp}{^{(\indice)}}
\newcommand{\angstrom}{\text{\normalfont\AA}}

% Time, effective population size and mutation rate.
\newcommand{\Ne}{N_{\mathrm{e}}}
\newcommand{\dnds}{\omega}
\newcommand{\Nsite}{\text{n}}
\newcommand{\site}{\text{i}}
% DNA

\newcommand{\ci}{\mathbb{S}_{t}}
\newcommand{\cj}{\mathbb{S}_{t}'}
\newcommand{\itoj}{\ci, \cj}
\newcommand{\setNeighbors}{\mathcal{M}\left(\ci\right)}
\newcommand{\setNonSynNeighbors}{\mathcal{N}\left(\ci\right)}
\newcommand{\setSynNeighbors}{\mathcal{S}\left(\ci\right)}
\newcommand{\submatrix}{q}

\newcommand{\phenoDiv}{S}
\newcommand{\s}{s}
\newcommand{\pheno}{x}
\newcommand{\dpheno}{\delta \pheno}
\newcommand{\phenoFold}{\Delta G}
\newcommand{\phenoFoldMin}{\alpha}
\newcommand{\dphenoFold}{\gamma}

@@

We provide a simple toy model, with $2$ parameters that shows a variety of behaviours. 
This model map the amino-acid sequence to the free energy (kcal/mol) between folded and unfolded state of the whole protein.
Each amino-acid site in the protein can be in the unstable or stable state.
A protein is optimal ($\phenoFold = \phenoFoldMin$) if all amino-acids are in the stable state.
Each amino-acid destabilizing the protein brings an increase of $\phenoFold$ equal to $\dphenoFold$.
\begin{equation}
\phenoFold = \phenoFoldMin + \dphenoFold * d, 
\end{equation}
where $d$ is the number of unstable amino-acid sites in the sequence, $\alpha$ and $\gamma$ are in kcal/mol.
Wrightian fitness is defined as the probability of our protein to be in the folded state, given by Boltzmann equation: 
\begin{equation}
w(\phenoFold) = \dfrac{1}{1 + e^{\beta \phenoFold }}, 
\end{equation}
where $\beta=1.686$ mol/kcal at room temperature.\\

In the original studies \cite{Williams2006, Goldstein2011, Pollock2012}, free energy of the folded state is computed using the $3$D structure of the folded state and pair-wise contact energies between neighboring amino-acid residues using Miya-zawa \& Jernigan contact potentials \cite{Miyazawa1985}.
The free energy distribution of unfolded states is approximated using $55$ decoy $3$D structures that supposedly represent a sample of possible unfolded states (see Supplementary Materials).\\

The parameters $\alpha$ and $\gamma$ of our toy model are set to match the average values observed in the original model.
$\alpha=-118$ kcal/mol, as it is the $\phenoFold$ of the most stable sequence of $300$ sites \cite{Goldstein2011}.
$\gamma=1.0$ kcal/mol, as the average observed increase in $\phenoFold$ per destabilizing mutation \cite{Zeldovich2007}.

\begin{figure*}[htb!]
\begin{mdframed}
	\centering
	\includegraphics[width=\textwidth] {artworks/XPopSizeYOmegaEq.pdf}
	\vspace*{0.2em}
	\caption{
		\textbf{$\bm{\dnds}$ response to change in $\bm{\Ne}$}.
		Panel A. Our toy model reproduces the observations of Goldstein \cite{Goldstein2013} that $\dnds$ at equilibrium is independent of $\Ne$. Indeed, whenever the parameters $\alpha$ and $\gamma$ are taken to match the average values observed in the original model, we observed no response of $\dnds$ to $\Ne$ (pink and brown lines).
		However, decreasing both $\alpha$ and $\gamma$ together implies a strong negative relationship between $\Ne$ and $\dnds$, as predicted under the nearly-neutral theory of evolution.
		Panel B. By reducing the number of sites of the protein, thus by effectively reducing the potential number of sites that can compensate for the change in $\Ne$, the $\dnds$ response to changes in $\Ne$ is stronger.
	}
	\label{fig:EqdndsNe}
\end{mdframed}
\end{figure*}

DNA sequence evolution are simulated under an origin-fixation model \cite{McCandlish2014}, where the rate of substitution of a new mutation with selection coefficient $s$ is:
\begin{equation}
p_{\text{fix}}(s) = \mu \dfrac{4 \Ne s}{1 + e^{- 4 \Ne s }}, 
\end{equation}
where $\mu$ is the mutation rate.\\

Our toy model reproduces the observations that $\dnds$ at equilibrium is independent of $\Ne$. Indeed, whenever the parameters $\alpha$ and $\gamma$ are taken to match the average values observed in the original model, we observed no response of $\dnds$ to $\Ne$.
However, decreasing both $\alpha$ and $\gamma$ together implies a strong negative relationship between $\Ne$ and $\dnds$, as predicted under the nearly-neutral theory of evolution.
Moreover, by reducing the number of sites of the protein, thus by effectively reducing the potential number of sites that can compensate for the change in $\Ne$, the $\dnds$ at equilibrium after changes in $\Ne$ is stronger \textit{Necessary mechanistic explanation for this fact} (see Figure \ref{fig:EqdndsNe}).

@@

Although the equilibrium value of $\dnds$ after changes in $\Ne$ is an important feature of the $\dnds$-$\Ne$ relationship, an other parameter that is often neglected is the relaxation time to reach the new equilibrium $\dnds$ \cite{Jones2016}.
\begin{figure*}[htb!]
\begin{mdframed}
	\centering
	\includegraphics[width=\textwidth]{artworks/SimuStabExponentialDecay.png}
	\vspace*{0.2em}
	\caption{
		\textbf{$\bm{\dnds}$ relaxation time to changes in $\bm{\Ne}$}.
		Panel A. Our result show that with low epistasis, the time taken to reach the new equilibrium value of $\dnds$ after a change in $\Ne$ is long, such that relaxation rate is on the order of the mutation rate.
		Panel B. With high epistasis, the relaxation time is short.
	}
	\label{fig:relaxStability}
\end{mdframed}
\end{figure*}
In the case of a fitness landscape with epistasis, after a change in $\Ne$, any site that is used to climb either up or down the fitness landscape will have diminishing return in other site of the sequences.
Not taking into account epistasis have the consequence of over-estimating the relaxation time to return to equilibrium of $\dnds$ after changes in $\Ne$ \ref{fig:relaxStability}).

@@

\begin{figure*}[htb!]
\begin{mdframed}
	\centering
	\includegraphics[width=\textwidth]{artworks/SimuExponentialDecay.png}
	\vspace*{0.2em}
	\caption{
		\textbf{$\bm{\dnds}$ relaxation time under different fitness models.}.
		Panel A. In context of a fixed fitness landscape, where each amino-acid have different fitnesses (site-specific profile), the time taken to reach the new equilibrium value of $\dnds$ after a change in $\Ne$ is long, such that relaxation rate is on the order of the mutation rate.
		Panel B. In the context of a distribution of fitness effects, the relaxation time is non-existent.
	}
	\label{fig:relaxProfileDFE}
\end{mdframed}
\end{figure*}

@@

\subsection{Simulation with protein folding probability}
\label{MatMet:folding}
From a DNA sequence $\ci$ after $t$ substitutions, the protein's difference in free energy between folded and unfolded state is given by:
\begin{equation*}
\phenoFold\left(\ci\right) = \phenoFoldMin + \dphenoFold * d\left(\ci\right), 
\end{equation*}
where $d\left(\ci\right)$ is the number of non-optimal amino-acid sites in the sequence.
Wrightian fitness is defined as the probability of our protein to be in the folded state, given by the Boltzmann equation: 
\begin{equation}
w(\phenoFold\left(\ci\right)) = \dfrac{P_{\mathrm{F}}\left(\ci\right)}{P_{\mathrm{F}}\left(\ci\right) + P_{\mathrm{U}}\left(\ci\right)} = \dfrac{e^{-\beta G_{\mathrm{F}}\left(\ci\right) }}{e^{-\beta G_{\mathrm{F}} \left(\ci\right) } + e^{-\beta G_{\mathrm{U}}\left(\ci\right) }} = \dfrac{1}{1 + e^{\beta \phenoFold\left(\ci\right) }}, 
\end{equation}
where $\beta$ is the inverse of the temperature ($\beta=1/kT$).

From the resident sequence $\ci$, we define $\setNeighbors$ as the set of all possible mutant that are one nucleotide away from $\ci$, and were mutant sequences containing a stop codon are excluded.
For a protein of $\Nsite$ amino-acid sites, $\left| \setNeighbors \right| \leq 9 \Nsite$, since each codon has a maximum of $9$ possible mutant codon that are one mutation away and that are not stop codon.
For each mutant sequence $\cj \in \setNeighbors$, we compute $\phenoFold^{t+1}$ from the updated sequence $\cj$, and subsequently the selection coefficient of the mutant:
\begin{equation}
s \left( \ci,\cj\right) = \dfrac{ w\left( \phenoFold\left(\cj\right) \right) - w\left( \phenoFold\left(\ci\right) \right)}{w\left( \phenoFold\left(\ci\right) \right)}.
\end{equation}
The next change in the protein coding DNA and the time to next the event is chosen using Gillespie algorithm, according to the rates of substitution between codons:
\begin{equation}
\submatrix_{\itoj} = \mu_{\itoj} \dfrac{4 \Ne s \left( \ci,\cj\right)}{1 - \e^{-4 \Ne s \left( \ci,\cj\right)}}, 
\end{equation}
where $\mu_{\itoj}$ is the mutation rate between $\ci$ and $\cj$, determined by the underlying $4x4$ nucleotide mutation rate matrix, and ${\submatrix_{\itoj}} = \mu_{\itoj}$ in the case of synonymous substitutions.

@@

\newpage
\part*{Supplementary materials}

\section*{Probability of folding as in Goldstein \& Pollock}
We simulated substitutions in the protein phosphatase ($\Nsite=300$ codon sites).
From a DNA sequence $\ci$ after $t$ substitutions, we compute the free energy of the folded state $G_{\mathrm{F}}\left(\ci\right)$, using the $3$-dimensional structure of the folded state and pair-wise contact energies between neighboring amino-acid residues:
\begin{equation}
G_{\mathrm{F}}\left(\ci\right) = \sum_{1 \leq \site \leq \Nsite} \sum_{r \in \mathcal{N}(\site)} I \left(\ci(\site), \ci(r) \right),
\end{equation}
where $I(a,b)$ is the pair-wise contact energies between amino-acid $a$ and $b$, using contact potentials estimated by Miya-zawa and Jernigan, and $\mathcal{N}(\site)$ are the neighbor residues of site $\site$ (closer than $7\angstrom$) in the $3$D structure.\\

The free energy of unfolded states $G_{\mathrm{U}}\left(\ci\right)$ is approximated using $55$ decoy $3$D structures that supposedly represent a sample of possible unfolded states:
\begin{equation}
G_{\mathrm{U}}\left(\ci\right) = \langle G\left(\ci\right) \rangle - kT \ln (1.0\mathrm{E}^{160}) - \dfrac{2 \left[ \langle G\left(\ci\right)^2 \rangle - \langle G\left(\ci\right) \rangle^2\right] }{kT}
\end{equation}
where the average $\langle . \rangle$ runs other the $55$ decoy $3$D structures, and $k$ is the Boltzmann constant and $T$ the temperature in Kelvin.\\

From the energy of folded and unfolded states, we can compute the difference in free energy between the states:
\begin{equation}
\phenoFold\left(\ci\right) = G_{\mathrm{F}}\left(\ci\right) - G_{\mathrm{U}}\left(\ci\right)
\end{equation}

\begin{center}
	\includegraphics[width=165mm] {artworks/ModelSimuFold.pdf}
\end{center}

\section*{Equimutability}
We define $\pheno$ as the current phenotype and $\dpheno$ as the phenotypic effect of a mutation.
$\s \left( \pheno, \dpheno \right)$ is the selection coefficient of a phenotypic effect $\dpheno$ if the current phenotype is $\pheno$:
\begin{equation*}
\s \left( \pheno, \dpheno \right) = \dfrac{w \left( \pheno + \dpheno \right) - w \left( \pheno \right)}{w \left( \pheno \right)} \simeq w'\left( \pheno \right) \dpheno, 
\end{equation*}
where $w \left( \pheno \right)$ is the Wrightian fitness of phenotype $x$.
We also define $\rho_{\pheno}(\dpheno)$ as the distribution of phenotypic effects, at phenotype $\pheno$.
At equilibrium, the expected selection coefficient $ 4 \Ne \s \left( \pheno, \dpheno \right)$ of mutations that reached fixation must be $0$:
\begin{align*}
0 & = \E_{\dpheno} \left[ 4 \Ne \s\left( \pheno, \dpheno \right) \dfrac{ 4 \Ne \s\left( \pheno, \dpheno \right) }{1 - \e^{4\Ne\s\left( \pheno, \dpheno \right)}} \right] \\ & 
\simeq \int_{-\infty}^{\infty} \dfrac{ \left[ 4 \Ne f'\left( \pheno \right)  \right]^2}{1 - \e^{4\Ne f'\left( \pheno \right) \dpheno} }  \dpheno^2  \rho_{\pheno}(\dpheno) \der \dpheno 
\end{align*}
And the substitution rate is :
\begin{align*}
\omega(\Ne) & = \E_{\dpheno} \left[ \dfrac{ 4 \Ne \s \left( \pheno, \dpheno \right)}{1 - \e^{4\Ne \s \left( \pheno, \dpheno \right)}} \right] \simeq \E_{\dpheno} \left[ \dfrac{ 4 \Ne f'\left( \pheno \right) \dpheno}{1 - \e^{4\Ne f'\left( \pheno \right) \dpheno}} \right] \\
& \simeq \int_{-\infty}^{\infty} \dfrac{4 \Ne f'\left( \pheno \right) \dpheno}{1 - \e^{4\Ne f'\left( \pheno \right) \dpheno } } \rho_{\pheno}(\dpheno) \der \dpheno \\
\end{align*}
If $\rho_{\pheno}(\dpheno) = \rho(\dpheno)$ is independent of $\pheno$:
\begin{align*}
0 & \simeq \int_{-\infty}^{\infty} \dfrac{ \left[ 4 \Ne f'\left( \pheno \right)  \right]^2}{1 - \e^{4\Ne f'\left( \pheno \right) \dpheno} } \dpheno^2 \rho(\dpheno)  \der \dpheno \\
\Longrightarrow & \ 4 \Ne f'\left( \pheno \right) = K \in \mathbb{R}^{*} \text{ and } \s\left( \pheno, \dpheno \right) \simeq \dfrac{K \dpheno}{4\Ne} \\
\Longrightarrow & \ \omega( \Ne ) \simeq \int_{-\infty}^{\infty} \dfrac{ K  }{1 - \e^{K \dpheno } } \dpheno \rho(\dpheno)  \der \dpheno  \text{ is independent of } \Ne
\end{align*}

>>>>

\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{adjustbox}
\usepackage{float}
\usepackage{caption}
\usepackage{mdframed}
\usepackage{lmodern}
\usepackage{bm,bbold}
\usepackage{xfrac, nicefrac}
\usepackage{xcolor}
\usepackage{lmodern}
\usepackage{enumitem}
\usepackage{pgfplots, pgf,tikz}
\usepackage[margin=60pt]{geometry}
\pdfinclusioncopyfonts=1
\captionsetup{width=0.85\textwidth}

\include{new-commands}
\pgfplotsset{every axis/.append style={line width=1pt}}
\pgfplotscreateplotcyclelist{colors}{LIGHTGREEN\\YELLOW\\RED\\GREEN\\BLUE\\}

@@

We provide a toy model of protein evolution under mutation-selection-drift equilibrium, reproducing the observation of more complex models, where we derive mathematical relationship using population-genetic equations. 
This model map the amino-acid sequence to the difference in free energy ($\deltaG$ in kcal/mol) between folded and unfolded state of the whole protein.
Each amino-acid site in the protein can be in the unstable or stable state.
\begin{equation}
\deltaG = \deltaGMin + \Nsite * \gamma * \x, 
\end{equation}
where $\x$ is the proportion of unstable amino-acid sites in the sequence, $\deltaGMin$ and $\gamma$ are in kcal/mol.
An optimal protein has a $\deltaG = \deltaGMin$, and pessimal protein has a $\deltaG = \deltaGMin + \Nsite \gamma$.
Wrightian fitness is defined as the probability of our protein to be in the folded state, given by Boltzmann equation: 
\begin{equation}
f(\deltaG) = \dfrac{1}{1 + e^{\beta \deltaG }}, 
\end{equation}
where the fixed parameter $\beta=1.686$ mol/kcal at room temperature.\\

We compare our model to the original studies \cite{Williams2006, Goldstein2011, Pollock2012}, where free energy of the folded state is computed using the $3$D structure of the folded state and pair-wise contact potential energies between neighboring amino-acid residues \cite{Miyazawa1985}.
The free energy distribution of unfolded states is approximated using $55$ decoy $3$D structures that supposedly represent a sample of possible unfolded states (see Supplementary Materials).

\begin{figure*}[htb!]
\begin{mdframed}
\centering
	\includegraphics[width=0.6\textwidth] {artworks/GoldsteinPollock/dnds_event_tot.pdf}\\
	\includegraphics[width=0.6\textwidth] {artworks/N300/dnds_event_tot.pdf}\\
	\includegraphics[width=0.6\textwidth] {artworks/N300Variance/dnds_event_tot.pdf}\\
	\caption{
		\textbf{$\bm{\dnds}$ response to change in $\bm{\Ne}$}.
		Top panel, $\dnds$ at equilibrium is weakly dependent on log-$\Ne$ in the model of Goldstein \& Pollock, but is not independent as claimed in \cite{Goldstein2013}.
		For each population size, $200$ simulation were performed and the average (blue solid line) and standard deviation (blue filling) are shown.
		This weak correlation matches the prediction of our toy model that the linear relation (blue dashed line) has a slope equal to $(\beta  \Nsite \gamma)^{-1} \simeq 0.0019$.
		Middle panel, $\alpha=-118$ and $\Nsite \gamma=300\E[\gamma]=300$.
		Bottom panel, $\alpha=-118$ and each site has it's own normally distributed $\gamma$ with mean $1$ and standard deviation of $1$.
		The original $3$D model of Goldstein \& Pollock and our simple toy model provide qualitatively and quantitatively the same linear relationship between $\dnds$ and log-$\Ne$, whenever the parameters $\alpha$ and $\Nsite \gamma$ are taken to match the average values observed in the original $3$D model.
		\label{fig:GoldsteinVsToy}
	}
\end{mdframed}
\end{figure*}

\begin{figure*}[htb!]
\begin{mdframed}
	\centering
	\includegraphics[width=0.8\textwidth] {artworks/ScaleAlphaGammaStdN300/SimuStab/dnds_event_tot.pdf}\\
	\caption{
		\textbf{$\bm{\dnds}$ linear response to change in $\bm{\Ne}$}.
		The parameter $\Nsite \gamma$ is controlling the $\dnds$-$\Ne$ relationship, decreasing $\Nsite \gamma$ implies a stronger negative linear relationship between $\dnds$ and log-$\Ne$. The intersect is a function of both $\alpha$ and $\Nsite \gamma$.
	}
	\label{fig:EqdndsNe}
\end{mdframed}
\end{figure*}

Protein sequence evolution are simulated under an origin-fixation model \cite{McCandlish2014}, where the rate of substitution of a new mutation with selection coefficient $s$ is:
\begin{equation}
p_{\text{fix}}(s) = \mu \dfrac{4 \Ne s}{1 + e^{- 4 \Ne s }}, 
\end{equation}
where $\mu$ is the mutation rate.\\

Simulations of protein evolution with both models display the same $\dnds$-$\Ne$ relationship, whenever the parameters $\alpha$ and $\Nsite \gamma$ of our model are set to match the average values of the $3$D model (see Figure \ref{fig:EqdndsNe}, panel A).
In such experiment, $\alpha=-118$ kcal/mol as it is the $\deltaG$ of the most stable sequence of $300$ sites \cite{Goldstein2011}.
The parameter $\Nsite \gamma=300.0$ kcal/mol, since the average observed increase in $\deltaG$ per destabilizing mutation is $\deltadeltaG = 1.0$ kcal/mol per site and our protein is $300$ sites \cite{Zeldovich2007}.
The original works claimed that under the $3$D model, $\dnds$ is independent of $\Ne$ because $\deltaG$ is an equimutable parameter.
We argue instead that $\dnds$ is approximatively linear with log-$\Ne$, and that $\deltaG$ is not equimutable (see Figure \ref{fig:EqdndsNe}, panel B).
More precisely, this mathematical equation can be derived (see Supplementary Materials) and the slope is inversely proportional to $\Nsite \gamma$ :
\begin{equation}
\dnds \propto -\dfrac{1}{\beta \Nsite \gamma} \ln(\Ne).
\end{equation}
Meaning that under a value of $\Nsite \gamma=300.0$ kcal/mol, the slope of the linear relationship between $\dnds$ and log-$\Ne$ is $0.002$.
Thus if we consider $\deltadeltaG = 1.0$ kcal/mol for destabilizing mutations, the slope is determined solely by the number of sites in the protein. 
However, decreasing both $\alpha$ and $\gamma$ together implies a strong negative relationship between $\Ne$ and $\dnds$, as predicted under the nearly-neutral theory of evolution.
Moreover, by reducing the number of sites of the protein, thus by effectively reducing the potential number of sites that can compensate for the change in $\Ne$, the $\dnds$ at equilibrium after changes in $\Ne$ is stronger \textit{Necessary mechanistic explanation for this fact} (see Figure \ref{fig:EqdndsNe}).

\begin{figure*}[htb!]
\begin{mdframed}
\centering
	\begin{tikzpicture}
	\begin{axis}[
	xlabel={$\x$},
	ymin=0.0, ymax=1.0,
	samples=100,
	legend entries={$\ln(\frac{1 - \x}{\x})$,
		$S_{\deltaGMin = -118,\Nsite \gamma = 300,\Ne = 10^{4}}(\x)$,
		$S_{\deltaGMin = -118,\Nsite \gamma = 300,\Ne = 10^{5}}(\x)$,
		$S_{\deltaGMin = -16.7,\Nsite \gamma = 30,\Ne = 10^{4}}(\x)$,
		$S_{\deltaGMin = -16.7,\Nsite \gamma = 30,\Ne = 10^{5}}(\x)$,
	},
	legend cell align=left,
	minor tick num=2,
	axis x line=bottom,
	axis y line=left,
	legend style={at={(0.5,0.95)},anchor=north west}
	]
	\addplot[domain=0.3:0.6,line width=2.0pt, color=BLUE]{ ln((1-x)/x) };
	\addplot[name path=A, domain=0.3:0.38, line width=2.0pt, color=RED]{4*10000*1.686*exp(1.686*(-118+300*x))};
	\addplot[name path=B, dashed,domain=0.3:0.37, line width=2.0pt, color=RED]{4*100000*1.686*exp(1.686*(-118+300*x))};
	\addplot[name path=C,domain=0.3:0.39, line width=2.0pt, color=GREEN]{4*10000*0.1*1.686*exp(1.686*(-16.71+30*x))};
	\addplot[name path=D,dashed,domain=0.3:0.35, line width=2.0pt, color=GREEN]{4*100000*0.1*1.686*exp(1.686*(-16.71+30*x))};
	\addplot[fill=RED, opacity=0.2] fill between[ of = A and B];
	\addplot[fill=GREEN, opacity=0.2] fill between[ of = C and D];
	\end{axis}
	\end{tikzpicture}
	\caption{
	\textbf{$\bm{\dnds}$ response to change in $\bm{\Ne}$}.
	The equilibrium $\x\eq$ is determined by the equation $\ln(\frac{1 - \x\eq}{\x\eq})=S_{\deltaGMin,\Nsite \gamma,\Ne}(\x\eq)$ (see Supplementary Materials). $S$ is proportional to $\Ne$, but increase exponentially with $\x$ where $\Nsite \gamma$ is the exponential growth rate.
	}
\label{fig:NeChangeInfluence}
\end{mdframed}
\end{figure*}

@@

Although the equilibrium value of $\dnds$ after changes in $\Ne$ is an important feature of the $\dnds$-$\Ne$ relationship, an other parameter that is often neglected is the relaxation time to reach the new equilibrium $\dnds$ \cite{Jones2016}.
\begin{figure*}[htb!]
\begin{mdframed}
	\centering
	\includegraphics[width=\textwidth]{artworks/RelaxationScaleAlphaGammaN100/14_1_dNdN0.png}
	\includegraphics[width=\textwidth]{artworks/RelaxationScaleAlphaGammaN100/100_1_dNdN0.png}\\
	\caption{
		\textbf{$\bm{\dnds}$ relaxation time to changes in $\bm{\Ne}$}.
		Panel A. Our result show that with low epistasis, the time taken to reach the new equilibrium value of $\dnds$ after a change in $\Ne$ is long, such that relaxation rate is on the order of the mutation rate.
		Panel B. With high epistasis, the relaxation time is short.
	}
	\label{fig:relaxStability}
\end{mdframed}
\end{figure*}
In the case of a fitness landscape with epistasis, after a change in $\Ne$, any site that is used to climb either up or down the fitness landscape will have diminishing return in other site of the sequences.
Not taking into account epistasis have the consequence of over-estimating the relaxation time to return to equilibrium of $\dnds$ after changes in $\Ne$ \ref{fig:relaxStability}).

@@

\begin{figure*}[htb!]
\begin{mdframed}
	\centering
	\includegraphics[width=\textwidth]{artworks/SimuExponentialDecay.png}\\
	\caption{
		\textbf{$\bm{\dnds}$ relaxation time under different fitness models.}.
		Panel A. In context of a fixed fitness landscape, where each amino-acid have different fitnesses (site-specific profile), the time taken to reach the new equilibrium value of $\dnds$ after a change in $\Ne$ is long, such that relaxation rate is on the order of the mutation rate.
		Panel B. In the context of a distribution of fitness effects, the relaxation time is non-existent.
	}
	\label{fig:relaxProfileDFE}
\end{mdframed}
\end{figure*}

@@

\subsection{Simulation with protein folding probability}
\label{MatMet:folding}
From a DNA sequence $\ci$ after $t$ substitutions, the protein's difference in free energy between folded and unfolded state is given by:
\begin{equation*}
\deltaG\left(\ci\right) = \deltaGMin + \Nsite \gamma * \x\left(\ci\right), 
\end{equation*}
where $\x\left(\ci\right)$ is the proportion of non-optimal amino-acid sites in the sequence.
Wrightian fitness is defined as the probability of our protein to be in the folded state, given by the Boltzmann equation: 
\begin{equation}
f(\deltaG\left(\ci\right)) = \dfrac{P_{\mathrm{F}}\left(\ci\right)}{P_{\mathrm{F}}\left(\ci\right) + P_{\mathrm{U}}\left(\ci\right)} = \dfrac{e^{-\beta G_{\mathrm{F}}\left(\ci\right) }}{e^{-\beta G_{\mathrm{F}} \left(\ci\right) } + e^{-\beta G_{\mathrm{U}}\left(\ci\right) }} = \dfrac{1}{1 + e^{\beta \deltaG\left(\ci\right) }}, 
\end{equation}
where $\beta$ is the inverse of the temperature ($\beta=1/kT$).

From the resident sequence $\ci$, we define $\setNeighbors$ as the set of all possible mutant that are one nucleotide away from $\ci$, and were mutant sequences containing a stop codon are excluded.
For a protein of $\Nsite$ amino-acid sites, $\left| \setNeighbors \right| \leq 9 \Nsite$, since each codon has a maximum of $9$ possible mutant codon that are one mutation away and that are not stop codon.
For each mutant sequence $\cj \in \setNeighbors$, we compute $\deltaG^{t+1}$ from the updated sequence $\cj$, and subsequently the selection coefficient of the mutant:
\begin{equation}
s \left( \ci,\cj\right) = \dfrac{ f\left( \deltaG \left(\cj\right) \right) - f\left( \deltaG \left(\ci\right) \right)}{f\left( \deltaG \left(\ci\right) \right)}.
\end{equation}
The next change in the protein coding DNA and the time to next the event is chosen using Gillespie algorithm, according to the rates of substitution between codons:
\begin{equation}
\submatrix_{\itoj} = \mu_{\itoj} \dfrac{4 \Ne s \left( \ci,\cj\right)}{1 - \e^{-4 \Ne s \left( \ci,\cj\right)}}, 
\end{equation}
where $\mu_{\itoj}$ is the mutation rate between $\ci$ and $\cj$, determined by the underlying $4x4$ nucleotide mutation rate matrix, and ${\submatrix_{\itoj}} = \mu_{\itoj}$ in the case of synonymous substitutions.

@@
